---
title: "A UN Report on the Relationship between GDP and Life Expectancy"
author: "Yumino"
date: '2023-01-18'
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-packages, include = FALSE}
library(tidyverse)
library(knitr)
```

This report was prepared to the attention of the United Nations. It analyzes the relationship between a country's GDP, its life expectancy and CO2 emissions. Our goal is to determine to what degree a country’s economic strength or weakness may be related to its public health status and impact on climate pollution. We hypothesize that both life expectancy and CO2 emissions increase with a country's GDP.

```{r read data, include = FALSE}
gapminder_1997 <- read_csv("data/gapminder_1997.csv")
```

```{r gdp_lifeExp_1997 plot, echo = FALSE, fig.width = 4, fig.align = "center"}
# make plot of GDP (x) vs life expectancy (y)
ggplot(data = gapminder_1997, aes(x = gdpPercap, y = lifeExp, color = continent)) +
        geom_point() +
        scale_color_brewer(palette = "YlGnBu") +
        labs(x = "GDP Per Capita", y = "Life Expectancy") +
        theme_bw()
```

```{r basic-country-stats}
# number of countries in dataset
nCountries <- gapminder_1997 %>%
        select(country) %>%
        n_distinct()

# minimum GDP
minGDP <- gapminder_1997 %>%
        summarize(round(min(gdpPercap))) %>%
        pull()

# maximum GDP
maxGDP <- gapminder_1997 %>%
        summarize(round(max(gdpPercap))) %>%
        pull()
```

The number of countries in the dataset is `r nCountries`.
The minimum GDP is `r minGDP`. The maximum GDP is `r maxGDP`.

```{r table}
gapminder_1997 %>%
        filter(continent == "Oceania") %>%
        kable()
```



